---
title: "Cross validation"
author: "Jinghan Zhao"
date: "2024-11-12"
output: github_document
---


```{r setup, include=FALSE}
library(tidyverse)
library(modelr)
library(mgcv)
library(SemiPar)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%",
  warning = FALSE
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

set.seed(1)
```

Look at LIDAR data

```{r}
data("lidar")

lidar_df = 
  lidar %>% 
  as_tibble() %>% 
  mutate(id = row_number())
```

```{r}
lidar_df %>% 
  ggplot(aes(x = range, y = logratio)) +
  geom_point()
```


## Try to do CV

We'll compare 3 models -- one linear, one smooth, one wiggly.

Construct training and testing df

```{r}
train_df = sample_frac(lidar_df, size = .8)
test_df = anti_join(lidar_df, train_df, by = "id")
```

Look at these

```{r}
ggplot(train_df, aes(x = range, y = logratio)) +
  geom_point() +
  geom_point(data = test_df, color = "red")
```

Fit three models (suppose to be a good model: not too simple, not too complex)

```{r}
linear_mod = lm(logratio ~ range, data = train_df)
smooth_mod = gam(logratio ~ s(range), data = train_df)
wiggly_mod = gam(logratio ~ s(range, k = 30), sp = 10e-6, data = train_df)
```

Look at fits

```{r}
train_df %>% 
  add_predictions(linear_mod) %>% 
  ggplot(aes(x = range, y = logratio)) +
  geom_point() +
  geom_point(data = test_df, color = "blue") +
  geom_line(aes(y = pred), color = "red")
```

Linear model is not complex enough

```{r}
train_df %>% 
  add_predictions(wiggly_mod) %>% 
  ggplot(aes(x = range, y = logratio)) +
  geom_point() +
  geom_line(aes(y = pred), color = "red")
```

Wiggly model is too complex: sensitive to slightly changes

```{r}
train_df %>% 
  add_predictions(smooth_mod) %>% 
  ggplot(aes(x = range, y = logratio)) +
  geom_point() +
  geom_line(aes(y = pred), color = "red")
```

Just right!


Compare these numerically using RMSE.

```{r}
rmse(linear_mod, test_df)
rmse(smooth_mod, test_df)
rmse(wiggly_mod, test_df)
```


## Repeat the train / test split
